---
title: "Westlake Stats"
author: "Charl Everts"
date: "28/07/2020"
output:
  html_document: default
  pdf_document: default
---
The data captured represents scores from 28 rounds played by Charl Everts at Westlake Golf Club. Scores are recorded from 2018 until present. As new rounds are recoreded, the scores are captured in the Westlake.xlsx file and read into R to update the data visualisations. The goal of this project is to identify trends and potential weaknesses in the user's scoring at Westlake, based on **Score, GIR and Putts per round.**

### Distribution of data

The following 5 graphs illustrate the distribution of:

- Front 9 scoring
- Back 9 scoring
- Gross scores
- GIR per round
- Putts per round

The graphs are followed by the *Six Number Summary* for each variable.

```{r, include = FALSE}
rm(list=ls())
library('rmarkdown')
setwd("C:/Users/user/Desktop/FinTech 2020")

# xls file
library(readxl)
library(ggplot2)
stats <- read_excel("Westlake.xlsx", sheet = "Tibble")
str(stats)
stats$`Front 9` <- as.integer(stats$`Front 9`)
stats$`Back 9` <- as.integer(stats$`Back 9`)
stats$`11`[21] <- 2
stats$`11`
colnames(stats)[which(names(stats) == "putts")] <- "Putts"
str(stats)
```



```{r, echo = FALSE, warning = FALSE}
library(ggplot2)
library(grid)
library(gridExtra)

# Front 9 histogram
F9 <- ggplot(stats, aes(x = `Front 9`)) + 
  geom_histogram(aes(y = ..density..), binwidth = 1, 
                 colour = "black", fill = "white") +
  geom_density(alpha = .2, fill = "#33CCFF") + 
  geom_vline(aes(xintercept = mean(na.omit(`Front 9`))),
             color = "deeppink", linetype = "dashed", size = 1) +
  scale_x_continuous(breaks = seq(32, 45, 1)) +
  ggtitle("Distribution of Front 9 scoring") +
  theme(plot.title = element_text(hjust = 0.5))

# Back 9 histogram
B9 <- ggplot(stats, aes(x = `Back 9`)) + 
  geom_histogram(aes(y = ..density..), binwidth = 1, 
                 colour = "black", fill = "white") +
  geom_density(alpha = .2, fill = "#33CCFF") + 
  geom_vline(aes(xintercept = mean(na.omit(`Back 9`))),
             color = "deeppink", linetype = "dashed", size = 1) +
  scale_x_continuous(breaks = seq(32, 45, 1)) +
  ggtitle("Distribution of Back 9 scoring") +
  theme(plot.title = element_text(hjust = 0.5))
```



```{r fig1, fig.height = 5, fig.width = 6, fig.align = "center", echo = FALSE}
F9
summary(stats$`Front 9`)
```



```{r fig2, fig.height = 5, fig.width = 6, fig.align = "center", echo = FALSE}
B9
summary(stats$`Back 9`)
```



```{r fig3, fig.height = 5, fig.width = 6, fig.align = "center", echo = FALSE}
# Gross histogram
Gross <- ggplot(stats, aes(x = Gross)) + 
  geom_histogram(aes(y = ..density..), binwidth = 1, 
                 colour = "black", fill = "white") +
  geom_density(alpha = .2, fill = "#33CCFF") + 
  geom_vline(aes(xintercept = mean(`Gross`)),
             color = "deeppink", linetype = "dashed", size = 1) +
  scale_x_continuous(breaks = seq(68, 85, 1)) +
  ggtitle("Distribution of Scores") +
  theme(plot.title = element_text(hjust = 0.5))
Gross
summary(stats$Gross)
```



```{r fig4, fig.height = 5, fig.width = 6, fig.align = "center", echo = FALSE}
# GIR histogram
GIR <- ggplot(stats, aes(x = GIR)) + 
  geom_histogram(aes(y = ..density..), binwidth = 1, 
                 colour = "black", fill = "white") +
  geom_density(alpha = .2, fill = "#33CCFF") + 
  geom_vline(aes(xintercept = mean(GIR)),
             color = "deeppink", linetype = "dashed", size = 1) +
  scale_x_continuous(breaks = seq(5, 18, 1)) +
  ggtitle("Distribution of GIR per round") +
  theme(plot.title = element_text(hjust = 0.5))
GIR
summary(stats$GIR)
```



```{r fig5, fig.height = 5, fig.width = 6, fig.align = "center", echo = FALSE}
# Putts histogram
Putts <- ggplot(stats, aes(x = Putts)) + 
  geom_histogram(aes(y = ..density..), binwidth = 1, 
                 colour = "black", fill = "white") +
  geom_density(alpha = .2, fill = "#33CCFF") + 
  geom_vline(aes(xintercept = mean(Putts)),
             color = "deeppink", linetype = "dashed", size = 1) +
  scale_x_continuous(breaks = seq(20, 42, 1)) +
  ggtitle("Distribution of Putts per round") +
  theme(plot.title = element_text(hjust = 0.5))
Putts
summary(stats$Putts)
```



```{r, include = FALSE}
# First Hole histogram
one <- ggplot(stats, aes(x = `1`)) + 
  geom_histogram(binwidth = 1, 
                 colour = "black", alpha = 0.5, fill = "#FF0000") +
  geom_vline(aes(xintercept = mean(`1`)),
             color = "black", linetype = "dashed", size = 1) +
  scale_y_continuous(breaks = seq(0, 20, 2))
one
summary(stats$`1`)

# Second Hole histogram
two <- ggplot(stats, aes(x = `2`)) + 
  geom_histogram(binwidth = 1, 
                 colour = "black", alpha = 0.5, fill = "#FF3333") +
  geom_vline(aes(xintercept = mean(`2`)),
             color = "black", linetype = "dashed", size = 1) +
  scale_y_continuous(breaks = seq(0, 20, 2))
two
summary(stats$`2`)

# Third Hole histogram
three <- ggplot(stats, aes(x = `3`)) + 
  geom_histogram(binwidth = 1, 
                 colour = "black", alpha = 0.5, fill = "#FF6666") +
  geom_vline(aes(xintercept = mean(`3`)),
             color = "black", linetype = "dashed", size = 1) +
  scale_y_continuous(breaks = seq(0, 20, 2))
three
summary(stats$`3`)

# Fourth Hole histogram
four <- ggplot(stats, aes(x = `4`)) + 
  geom_histogram(binwidth = 1, 
                 colour = "black", alpha = 0.5, fill = "#FF3300") +
  geom_vline(aes(xintercept = mean(`4`)),
             color = "black", linetype = "dashed", size = 1) +
  scale_y_continuous(breaks = seq(0, 20, 2))
four
summary(stats$`4`)

# Fifth Hole histogram
five <- ggplot(stats, aes(x = `5`)) + 
  geom_histogram(binwidth = 1, 
                 colour = "black", alpha = 0.5, fill = "#FF6600") +
  geom_vline(aes(xintercept = mean(`5`)),
             color = "black", linetype = "dashed", size = 1) +
  scale_y_continuous(breaks = seq(0, 20, 2))
five
summary(stats$`5`)

# Sixth Hole histogram
six <- ggplot(stats, aes(x = `6`)) + 
  geom_histogram(binwidth = 1, 
                 colour = "black", alpha = 0.5, fill = "#FF9900") +
  geom_vline(aes(xintercept = mean(`6`)),
             color = "black", linetype = "dashed", size = 1) +
  scale_y_continuous(breaks = seq(0, 20, 2))
six
summary(stats$`6`)

# Seven Hole histogram
seven <- ggplot(stats, aes(x = `7`)) + 
  geom_histogram(binwidth = 1, 
                 colour = "black", alpha = 0.5, fill = "#FFCC00") +
  geom_vline(aes(xintercept = mean(`7`)),
             color = "black", linetype = "dashed", size = 1) +
  scale_y_continuous(breaks = seq(0, 20, 2))
seven
summary(stats$`7`)

# Eighth Hole histogram
eight <- ggplot(stats, aes(x = `8`)) + 
  geom_histogram(binwidth = 1, 
                 colour = "black", alpha = 0.5, fill = "#FFFF00") +
  geom_vline(aes(xintercept = mean(`8`)),
             color = "black", linetype = "dashed", size = 1) +
  scale_y_continuous(breaks = seq(0, 20, 2))
eight
summary(stats$`8`)

# Ninth Hole histogram
nine <- ggplot(stats, aes(x = `9`)) + 
  geom_histogram(binwidth = 1, 
                 colour = "black", alpha = 0.5, fill = "#FFFF66") +
  geom_vline(aes(xintercept = mean(`9`)),
             color = "black", linetype = "dashed", size = 1) +
  scale_y_continuous(breaks = seq(0, 25, 2))
nine
summary(stats$`9`)
```

### Hole By Hole Scoring

Each of the 18 histogram graphs detail the frequency of scores on each hole. The mean score relative to par on each hole is marked by a dashed vertical line. The designations (x-axis) have the following meanings:

- -2: eagle
- -1: birdie
- 0: par
- 1: bogey
- 2: double bogey

```{r fig6, fig.height = 2, fig.width = 6, fig.align = "center", echo = FALSE}
grid.arrange(grobs = list(one, two, three), 
             nrow = 1, main = "Front Nine")
grid.arrange(grobs = list(four, five, six), 
             nrow = 1, main = "Front Nine")
grid.arrange(grobs = list(seven, eight, nine), 
             nrow = 1, main = "Front Nine")
```


```{r, include = FALSE}
# Tenth Hole histogram
ten <- ggplot(stats, aes(x = `10`)) + 
  geom_histogram(binwidth = 1, 
                 colour = "black", alpha = 0.5, fill = "#CCFF00") +
  geom_vline(aes(xintercept = mean(`10`)),
             color = "black", linetype = "dashed", size = 1) +
  scale_y_continuous(breaks = seq(0, 20, 2))
ten
summary(stats$`10`)

# Eleventh Hole histogram
eleven <- ggplot(stats, aes(x = `11`)) + 
  geom_histogram(binwidth = 1, 
                 colour = "black", alpha = 0.5, fill = "#99FF33") +
  geom_vline(aes(xintercept = mean(`11`)),
             color = "black", linetype = "dashed", size = 1) +
  scale_y_continuous(breaks = seq(0, 20, 2))
eleven
summary(stats$`11`)

# Twelfth Hole histogram
twelve <- ggplot(stats, aes(x = `12`)) + 
  geom_histogram(binwidth = 1, 
                 colour = "black", alpha = 0.5, fill = "#00FF00") +
  geom_vline(aes(xintercept = mean(`12`)),
             color = "black", linetype = "dashed", size = 1) +
  scale_y_continuous(breaks = seq(0, 20, 2))
twelve
summary(stats$`12`)

# Thirteenth Hole histogram
thirteen <- ggplot(stats, aes(x = `13`)) + 
  geom_histogram(binwidth = 1, 
                 colour = "black", alpha = 0.5, fill = "#66FFCC") +
  geom_vline(aes(xintercept = mean(`13`)),
             color = "black", linetype = "dashed", size = 1) +
  scale_y_continuous(breaks = seq(0, 20, 2))
thirteen
summary(stats$`13`)

# Fourteenth Hole histogram
fourteen <- ggplot(stats, aes(x = `14`)) + 
  geom_histogram(binwidth = 1, 
                 colour = "black", alpha = 0.5, fill = "#66FFFF") +
  geom_vline(aes(xintercept = mean(`14`)),
             color = "black", linetype = "dashed", size = 1) +
  scale_y_continuous(breaks = seq(0, 20, 2))
fourteen
summary(stats$`14`)

# Fifteenth Hole histogram
fifteen <- ggplot(stats, aes(x = `15`)) + 
  geom_histogram(binwidth = 1, 
                 colour = "black", alpha = 0.5, fill = "#00FFFF") +
  geom_vline(aes(xintercept = mean(`15`)),
             color = "black", linetype = "dashed", size = 1) +
  scale_y_continuous(breaks = seq(0, 20, 2))
fifteen
summary(stats$`15`)

# Sixteenth Hole histogram
sixteen <- ggplot(stats, aes(x = `16`)) + 
  geom_histogram(binwidth = 1, 
                 colour = "black", alpha = 0.5, fill = "#33CCFF") +
  geom_vline(aes(xintercept = mean(`16`)),
             color = "black", linetype = "dashed", size = 1) +
  scale_y_continuous(breaks = seq(0, 20, 2))
sixteen
summary(stats$`16`)

# Seventeenth Hole histogram
seventeen <- ggplot(stats, aes(x = `17`)) + 
  geom_histogram(binwidth = 1, 
                 colour = "black", alpha = 0.5, fill = "#0066FF") +
  geom_vline(aes(xintercept = mean(`17`)),
             color = "black", linetype = "dashed", size = 1) +
  scale_y_continuous(breaks = seq(0, 20, 2))
seventeen
summary(stats$`17`)

# Eighteenth Hole histogram
eighteen <- ggplot(stats, aes(x = `18`)) + 
  geom_histogram(binwidth = 1, 
                 colour = "black", alpha = 0.5, fill = "#0000FF") +
  geom_vline(aes(xintercept = mean(`18`)),
             color = "black", linetype = "dashed", size = 1) +
  scale_y_continuous(breaks = seq(0, 20, 2))
eighteen
summary(stats$`18`)
```


```{r fig7, fig.height = 2, fig.width = 6, fig.align = "center", echo = FALSE}
grid.arrange(grobs = list(ten, eleven, twelve), 
             nrow = 1, main = "Back Nine")
grid.arrange(grobs = list(thirteen, fourteen, fifteen), 
             nrow = 1, main = "Back Nine")
grid.arrange(grobs = list(sixteen, seventeen, eighteen), 
             nrow = 1, main = "Back Nine")
```


### The effect of GIR per round on Scoring
A **weak negative** relationship exists between GIR and Score. This is expected, since an increase in GIR hit generally leads to lower scoring, and vice versa.

```{r fig8, fig.height = 5, fig.width = 8, fig.align = "center", echo = FALSE, message = FALSE}
p <- ggplot(stats, aes(GIR, Gross))
p + geom_point(colour = "black", size = 3) +
  scale_x_continuous(breaks = seq(0, 18, 1)) +
  scale_y_continuous(breaks = seq(68, 82, 1)) +
  geom_smooth(method = "lm", alpha = .5) +
  ggtitle("GIR vs Scoring: Line of Best Fit") +
  theme(plot.title = element_text(hjust = 0.5))
```


### The effect of Putts per round on Scoring
A **strong positive** relationship exists between Putts and Score. This agrees with intuitition, since a worse putting stat (more putts per round) is associated with higher scoring patterns. It is clear that a combination of maximising GIR and minimising Putts per round will lead to a decrease in Scoring.

```{r fig9, fig.height = 5, fig.width = 8, fig.align = "center", echo = FALSE, message = FALSE}
p <- ggplot(stats, aes(Putts, Gross))
p + geom_point(colour = "black", size = 3) +
  scale_x_continuous(breaks = seq(20, 42, 1)) +
  scale_y_continuous(breaks = seq(68, 82, 1)) +
  geom_smooth(method = "lm", alpha = .5) +
  ggtitle("Putts vs Scoring: Line of Best Fit") +
  theme(plot.title = element_text(hjust = 0.5))
```


### Multiple Linear Regression
In order to substantiate the claims made above, a *multiple linear regression* is conducted. GIR and Putts per round are regressed on Gross Score in order to evaluate the statistical significance of the predictor variables on the response: Gross Score.

Since the *p-value is less than 0.05 for GIR and Putts*, it is concluded that both GIR and Putts per round are **statistitically significant** in predicting Gross score. They both have considerable predictive power for the response variable, and should therefore be included in the multiple linear regression model.

```{r, include = FALSE}
stats$`Front 9` <- as.numeric(stats$`Front 9`)
stats$`Back 9` <- as.numeric(stats$`Back 9`)
```

```{r, include = FALSE}

model1 <- lm(Gross ~ GIR + Putts, data = stats) 
summary(model1)   
```

```{r, echo = FALSE, warning = FALSE}
library(broom)
library(pander)
table_obj <- tidy(model1)      
pander(table_obj, digits = 3)   
```
 
### Interpretation of Model

The multiple linear regression model takes the following form:
$$f(X) = \beta_0 + \beta_1X_1 + \beta_2X_2$$
Where $f(X)$ is the response variable Gross Score and $X_i$ refers to the predictor variables: $X_1$: GIR & $X_2$: Putts.  
$\beta_i$ refers to the coefficient of the predictor variables: $\beta_1$: GIR & $\beta_2$: Putts.  

Using the coefficients determined by the model, the following relationship between predictor and response variables are established:

$$ Gross Score = 59.8 - 0.822*GIR + 0.787*Putts$$

The following example illustrates the above equation in practice.
Suppose you hit 12 GIR and have 26 putts for your round, your expected Gross Score is calculated as follows:
$$ Gross Score = 59.8 - 0.822*12 + 0.787*26$$
$$ Gross Score = 70.398$$

```{r, include = FALSE}
to_par <- c(sum(stats$`1`), sum(stats$`2`), sum(stats$`3`), 
       sum(stats$`4`), sum(stats$`5`), sum(stats$`6`), sum(stats$`7`),
       sum(stats$`8`), sum(stats$`9`), sum(stats$`10`), sum(stats$`11`),
       sum(stats$`12`), sum(stats$`13`), sum(stats$`14`), sum(stats$`15`),
       sum(stats$`16`), sum(stats$`17`), sum(stats$`18`))
to_par

average_hole <- c(mean(stats$`1`), mean(stats$`2`), mean(stats$`3`), 
                  mean(stats$`4`), mean(stats$`5`), mean(stats$`6`), 
                  mean(stats$`7`), mean(stats$`8`), mean(stats$`9`), 
                  mean(stats$`10`), mean(stats$`11`), mean(stats$`12`),
                  mean(stats$`13`), mean(stats$`14`), mean(stats$`15`),
                  mean(stats$`16`), mean(stats$`17`), mean(stats$`18`))
average_hole <- round(average_hole, 3)

ecclectic <- c(min(stats$`1`), min(stats$`2`), min(stats$`3`), 
               min(stats$`4`), min(stats$`5`), min(stats$`6`), min(stats$`7`),
               min(stats$`8`), min(stats$`9`), min(stats$`10`), min(stats$`11`),
               min(stats$`12`), min(stats$`13`), min(stats$`14`), min(stats$`15`),
               min(stats$`16`), min(stats$`17`), min(stats$`18`))
ecclectic
```

```{r, include = FALSE}
hole <- (1:18)
hole
```

```{r, include = FALSE}
stroke_rating <- as.data.frame(cbind(hole, to_par))
stroke_rating <- as.data.frame(cbind(stroke_rating, average_hole))
stroke_rating <- as.data.frame(cbind(stroke_rating, ecclectic))
stroke_rating
```

### Scoring Skyline
This graph is coloured according to cumulative score to par in all recorded rounds.

```{r fig10, fig.height = 7, fig.width = 10, fig.align = "center", echo = FALSE}
p <- ggplot(stroke_rating, aes(x = hole, y = to_par, fill = to_par)) +
  geom_bar(stat = "identity") + theme_minimal() + 
  scale_fill_gradient(low = "green", high = "red") +
  scale_x_discrete(limits = c(1:18)) +
  ggtitle("Scoring relative to par by hole") +
  theme(plot.title = element_text(hjust = 0.5)) +
  ylab("Relative to par") + xlab("Hole") +
  scale_y_continuous(breaks = seq(-10, 20, 1)) 
p 
```

The following table summarises the effective stroke rating of each hole based on the cumulative score to par for all recorded rounds.

```{r, include = FALSE}
library(broom)
library(pander)
stroke_rating <- as.data.frame(cbind(hole, to_par))
stroke_rating <- as.data.frame(cbind(stroke_rating, average_hole))
stroke_rating <- as.data.frame(cbind(stroke_rating, ecclectic))
pander(stroke_rating, digits = 3) 
```


```{r, include = FALSE}
stroke_rating <- stroke_rating[order(stroke_rating$to_par),]
stroke_rating
stroke <- (18:1)
stroke
stroke_rating <- cbind(stroke,stroke_rating)
row.names(stroke_rating) <- NULL
stroke_rating
# Ecclectic score
sum(ecclectic) + 72
```

```{r, echo = FALSE, warning = FALSE}
library(broom)
library(pander)
pander(stroke_rating, digits = 3) 
```

### Box and whisker plots of Front9/Back9/Putts/Gross/GIR
```{r fig11, fig.height = 3, fig.width = 7, fig.align = "center", echo = FALSE}
par(mfrow = c(1,1))
boxplot(stats$`Front 9`, col = "magenta", horizontal = T, 
        main = "Front 9 ")
boxplot(stats$`Back 9`, col = "magenta", horizontal = T, 
        main = "Back 9 ")
boxplot(stats$Putts, col = "magenta", horizontal = T, 
        main = "Putts")
boxplot(stats$Gross, col = "magenta", horizontal = T, 
        main = "Gross")
boxplot(stats$GIR, col = "magenta", horizontal = T, 
        main = "GIR")
```